---
title: "Data Camp / ARIMA Modeling with R - Chapter 4"
output:
  pdf_document: default
  html_notebook: default
---

Initialize and load the data

```{r}
library(astsa)
library(xts)
```

Chapter 4 of [Data Camp - ARIMA Modeling with R](https://www.datacamp.com/courses/arima-modeling-with-r)

## Seasonal ARIMA models

The letter s denotes the seasonal period.  i.e. 4 quarters, 12 months, 1.203849928849390 nano-seconds.

### Pure Seasonal Models

consider pure seasonal models such as an $SAR(P=1)_{1=12}$ model.  This would be something like monthly temperatures where the temperature for the current period is dependant on the one from 12 months prior + some noise.

$X_t = \phi X_{t-12} + W_t$

### ACF and PACF of Pure Seasonal Models

ACF and PACF behave as they do with the non-seasonal but only at the seasonal points, i.e. for S=4, 0, 4, 8, 12, ... or for s=12, 0, 12, 24, ...

| $SAR(P)_s$ | $SMA(Q)_s$ | $SARMA(P,Q)_s$
----|----|----|----
ACF* | Tails off | Cuts off lag QS | Tails off
PACF* | Cuts off lag PS|Tails off| Tails off

* The values at the nonseasonal lags are 0

### Mixed Seasonal Model

These are called: $\text{SARIMA}(p,d,q) x (P,D,Q)_s$ models.

The **p,d,q** are the orders of the non-seasonal part of the model and the upper case letters the seasonal part.

Consider $SARIMA(0,0,1)x(1,0,0)_12$ - $X_t = \phi X_{t-12} + W_t + \theta W_{t-1}$

This model says this months value is dependant on 

* **SAR(1)**: The value from 12 periods ago $X_{t-12}$
* **MA(1)**: The _shock_ (error) from the previos period.  $W_{t-1}$

## Determining what type of Mixed Model.

Looking at the P/ACF

1. Focus first on the seasonal part (lags 12, 24, ...) to see what the P/ACF values are doing to determine the model(s) to try.
2. Then focus on the non-seasonal part (lags 1, 2, 3, ...) to determine the best model to choose.

## Analysis of AirPassengers

* Data has trend and non-linear growth, that is the variance is increasing..
* log the data, which removes the non-linear growth and controls the values
* diff the data to remove the trend.

Note that there is what appears to be seasonality every 12 periods (months), this can be removed with a call to `diff(d, lag=12)` which will diff the data with lag=12.

* diff the data with lag=12 to remove the 12 period seasonality.

```{r}
par(mfrow=c(2,2))
plot(AirPassengers)
plot(log(AirPassengers))
plot(diff(log(AirPassengers)))
plot(diff(diff(log(AirPassengers)),12))
```

At thsi point we have d=1, D=1 and S=12.

Apply P/ACF the diffed data.  max.lag=48 for 48 periods (4 years)

ACF appears to cut off at s = 12 adn PACF appears to be trailing off which suggests a SMA, so P=0 and Q=1.

ACF and PACF appear to be trailing off which implies an arma so p=1 q=1

So we conclude an SARIMA(1,1,1)(0,1,1) on the log data.  However, the sarima shows no significance for the ar1 component, so we adjust the model to SARIMA(0,1,1)(0,1,1)

```{r}
acf2(diff(diff(log(AirPassengers)),12), max.lag = 48)
```

```{r}

fit = sarima(log(AirPassengers), 1, 1, 1, 0, 1, 1, 12)
fit$ttable

fit1 = sarima(log(AirPassengers), 0, 1, 1, 0, 1, 1, 12)
fit1$ttable
```

## Let's do it again with seasonal unemployment

Data appears to have a trend.  Difference the data to remove the trend.

Data appears to have seasonality.  Difference again with lag=12 (we could look closely to see this and it's there.)

Finally, no trend or seasonality, just WN.



```{r}
par(mfrow=c(2,2))
plot(unemp)
du <- diff(unemp)
plot(du)
ddu <- diff(du, lag=12)
plot(ddu)

```

## P/ACF the ddu data

Set lag to 60 for 5 years.

ACF at seasonal intervals drops off and PACF tails off, so SMA(1)

ACF at regular intervals tails off at lag=2 and seaonsl drops off at lag=2, so MA(2)

We now have SARIMA(0, 1, 1)(1, 1, 0) S=12

```{r}
acf2(ddu, max.lag=60)
```

And **_BAM!_** well fitting model!  Me know something!

```{r}
fit = sarima(unemp, 2, 1, 0, 0, 1, 1, 12)
fit$ttable
```

```{r}
par(mfrow=c(2,2))

plot(chicken)
plot(diff(chicken))
plot(diff(diff(chicken), 12))

```

There is seasonal persistence, the fact that the ACF values at seasonal intervals persists.  So, you need to 
difference it again.

```{r}
acf2(diff(chicken), max.lag=60)
```

```{r}
acf2(diff(diff(chicken), 12), max.lag=60)
```

## Births

```{r}
par(mfrow=c(2,1))
plot(birth)
db = diff(birth)
plot(db)
```

```{r}
acf2(db, max.lag = 60)
```

There is seasonal persistence, lines not going away at seasonal lags, so difference again but with lag=12.

```{r}
ddb = diff(db, lag=12)
acf2(ddb, max.lag=60)
```

Seasonal ACF drops and PACF trails off at seasonal lag 1, so SMA(1).

ACF drops at lag=1 and PACF trails off at lag=1.

Had to diff 2 x so model is 0, 1, 1, 0, 1, 1, 12?

```{r}
sarima(birth, 0, 1, 1, 0, 1, 1, 12)
```

Residual analysis isn't ideal.  Perhaps the drop in ACF was a trail off, try adding an AR

```{r}
sarima(birth, 1, 1, 1, 0, 1, 1, 12)
```

AR coefficient is significant and residuals are looking better.

## Forecasting.

Like with simpla ARIMA forecasting is simple.  **astsa** includes an `sarima.for()` function for forecasting seasona arimas.

Let's use the unemployment example from previously.  This was modeled as **SARIMA(2, 1, 0, 0, 1, 1, 12)**

```{r}
sarima.for(unemp, 2, 1, 0, 0, 1, 1, 12, n.ahead=36)
```

And the chicken data, which was modeled to **SARIMA(2, 1, 0, 1, 0, 0, 12)**

```{r}
sarima.for(chicken, 2, 1, 0, 1, 0, 0, 12, n.ahead=60)
```