---
title: "Data Camp / xts - Chapter 3"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

Chapter 3 of Data Camp xts class excersizes

## Merging data

### merge() / cbind()

`merge()` & `cbind()` merge xts by columns

Like data base joins but on dates.

Inner, outer, left and right joins.

Defaults is an "outer" join:

`merge(x, y)`

merges all rows of x and y and fills in missing values with NA.

`merge(x, y, join='inner')`

merges only rows where both data objects have data for the same dates.

`merge(x, y, join='right', fill=na.locf)`

merges rows based on dates in the right side (y in this example).  Missing values for x will be filled with NA,
except that we specified the fill=na.locf, which tells it to fill with the last records value.

Additionally merging a 'c' vector adds it as a new column.

`merge(x, c(2, 3, 4))`

and adding a scalar value will add a new column as well.

`merge(x, 5)`

**Normal R recycling rule apply.**

Adding a vector of time objects is also possible.  This will add missing dates in the series, initializing to NA or whatever method is indicated by the _fill=_ argument.

### rbind()

* Combine weries by **row**
* Rows inserted in tim order SO ...
* All rows in `rbind()` must have a time
* The number of columns must match

**NOTE**: The result may have duplicate time indexes if both objects have values with the same time index.

#### Some merge examples

```{r}
library(xts)

a <- as.xts(read.table(header=TRUE, text="
                    a
2016-06-05 -1.2070657
2016-06-08  0.2774292
2016-06-09  1.0844412
2016-06-13 -2.3456977"))

b <- as.xts(read.table(header=TRUE, text="
                    b
2016-06-05  0.4291247
2016-06-06  0.5060559
2016-06-08 -0.5747400
2016-06-09 -0.5466319"))

merge(a, b, join='inner')

merge(a, b, join='left', fill=0)
```

### Some rbind examples

```{r}
temps <- as.xts(read.table(header=TRUE, text="
           Temp.Max Temp.Mean Temp.Min
2016-07-01       74        69       60
2016-07-02       78        66       56
2016-07-03       79        68       59
2016-07-04       80        76       69
2016-07-05       90        79       68
2016-07-06       89        79       70
2016-07-07       87        78       72
2016-07-08       89        80       72
2016-07-09       81        73       67
2016-07-10       83        72       64
2016-07-11       93        81       69
2016-07-12       89        82       77
2016-07-13       86        78       68
2016-07-14       89        80       68
2016-07-15       75        72       60
2016-07-16       79        69       60"))

temps_june30 <- as.xts(read.table(header=TRUE, text="
           temp_max temp_mean temp_min
2016-06-30       75        73       63"))

temps_july17 <- as.xts(read.table(header=TRUE, text="
           temp_max temp_mean temp_min
2016-07-17       79        70       68"))

temps_july18 <- as.xts(read.table(header=TRUE, text="
           temp_max temp_mean temp_min
2016-07-18       75        70       65"))

# Row bind temps_june30 to temps, assign this to temps2
temps2 <- rbind(temps, temps_june30)

# Row bind temps_july17 and temps_july18 to temps2, call this temps3
temps3 <- rbind(temps2, temps_july17, temps_july18)

print(temps2)
print(temps3)
```

### Handling Missingness

xts inherits the zoo missingness functionality.

* l.o.c.f "last observation caried forward" - Fill NA's with last observations values.

> na.locf(object,
          na.rm=TRUE,
          fromLast=FALSE,
          maxgap=Inf)
          
na.rm    - Should leading NA's be removed?
fromLast - Should the value from the next value obs be used instead of previous.  n.o.c.f "next obs...."
maxgap   - How many to replace before stopping

#### Fill all NAs

* Replace NA's

> na.fill(object, fill, ...)

* Remove NA's

> na.trim(object, ...)
> na.omit(object, ...)

* Interpolate NA's

The use of these should be carefully considered since they look forward in the data and could introduce look-ahead bias into the data which can affect forecasting.

> na.approx(object, ...)
> na.split(object, ...)

### Example

```{r}
temps <- as.xts(read.table(header=TRUE, text="
           Temp.Max Temp.Mean Temp.Min
2016-07-01       74        69       60
2016-07-02       78        NA       56
2016-07-03       79        NA       59
2016-07-04       80        76       69
2016-07-05       90        79       68"))

temps_last <- na.locf(temps)

temps_next <- na.locf(temps, fromLast=TRUE)

print(temps_last)
print(temps_next)
```

#### na.approx()

Let's look at an example of na.approx using the AirPassengers data.

1. First make it an xts
2. Sub-set the data to 1960
2. duplicate the data columnn replacing some of the data with NA's

```{r}
AirPass <- as.xts(AirPassengers)['1960']
indexFormat(AirPass) <- '%Y-%m-%d'
nr <- coredata(AirPass)
nr[c(3,4,5)] <- NA

AirPass <- cbind(AirPass, nr)
names(AirPass) <- c('original', 'missing')
print(AirPass)

na.approx(AirPass)
na.spline((AirPass))
```

## Seasonality and stationarity

### lag  (backshift operation)

What if you want to compare today's value with yesterdays?

`lag()` operator will shift times to shift times back, or forward.

> lag(x, 
      k=1, 
      na.pad=TRUE, ...)

* k controls number of lags
* na.pad controls NA introduction


#### Differences with zoo and ts lag()

* Positive k is forward, negative k is backward.  zoo and ts are the opposite.
* How missingness is handled after the shift

k can be a vector of values which will cause a new matrix with 1 column for each lag specified in the vector.

### diff

Calculate the difference of a lag.

> diff(x,
      lag=1,
      differences=1,
      arithmetic=TRUE,
      log=FALSE,
      na.pad=TRUE, ...)
      
      
### Example of lag()

```{r}
x <- as.xts(read.table(header=TRUE, text="
           x
2017-05-22 5
2017-05-23 6
2017-05-24 7
2017-05-25 8
2017-05-26 9"))

# "lead of x" means go back in time
lead_x <- lag(x, k=(-1))

# "lag of x" means go forward in time
lag_x <- lag(x, k=1)

z <- merge(lead_x, x, lag_x)

print(lead_x)
print(lag_x)
print(z)
```
      
### example of diff()

```{r}
AirPass <- as.xts(AirPassengers)
indexFormat(AirPass) <- '%Y-%m-%d'
names(AirPass) <- c('passengers')

diff_by_hand <- AirPass - lag(AirPass, k=1)

cbind(head(diff_by_hand), head(diff(AirPass)))

head(cbind(AirPass, diff(AirPass, lag = 12, differences = 1)), n=36)
```


