---
title: "Data Camp / xts - Chapter 2"
output:
  pdf_document: default
  html_notebook: default
---

Chapter 2 of Data Camp xts class excersizes

xts uses the [ISO-8601 standard](https://en.wikipedia.org/wiki/ISO_8601) for time storage.  The standard defines
time formats as well as ranges.

It is **important** to follow the standard exactly when indexing on dates.  The [subset](https://www.rdocumentation.org/packages/xts/versions/0.9-7/topics/%5B.xts) and [.parseISO8601](https://www.rdocumentation.org/packages/xts/versions/0.9-7/topics/.parseISO8601) documentation provies more details.

For full details see the [ISO page for ISO-8601](https://www.iso.org/iso-8601-date-and-time-format.html)

#### Some subsetting examples

```{r}
library(xts)

dates <- seq.Date(as.Date('2015-06-01'), 
                  as.Date('2016-04-01'),
                  by='day')

x <- xts(x=rnorm(length(dates)), 
         order.by=dates)

x_2016 <- x['2016']

jan_march <- x['2016/20160322']

print(str(x_2016))
print(str(jan_march))

82 == length(jan_march)
```

### Date + Time info

```{r}

date_times <- seq(as.POSIXct('2010-01-01T00:00:00'), as.POSIXct('2010-12-31T23:59:00'), by='hour')
str(date_times)

irreg <- xts(x=runif(length(date_times)),
             order.by=date_times)

morn_2010 <- irreg['T08/T10:00']
str(morn_2010)

morn_2010['2010-01-13']
```

### Alternative methods of extraction

_matrix_ like integer vectors  
_matrix_ like logical vectors  
Using POSIX or Date types (not common, or recommended)  

### Modifying data
Can use all supported indexing methods to update values in the xts object.

Also, **which.i=* option, which, liek the `which()` function, retuns the indexes of the rows instead of a matrix.

### Key behaviors

* All subsets preserve matrix (drop = FALSE).  xts always returns a matrix.
* Order is preserved.
* Binary search and memcpy are faster than base R!
* index and xts attributes are generally preserved.

## Example of replacement

```{r}
x <- xts(0:50,
         as.Date('2016-03-01') + 0:50)

dates <- as.Date('2016-03-10') + 0:5

x[dates] <- NA

print(x['20160308/17'])

x['20160401/'] <- 0
print(x['20160320/'])
```

### Finding times of interest

* R uses `head()` & `tail()` to look at the start and end of a series
* xts implements 2 similar with respect to time
  * uses a flexible notion of time
  * i.e. "last 3 days" or "first 6 weeks"
* These are the `first()` and `last()` functions

value can be soemthing like:

* "4 months", " 1 year", etc.
* can use numeric values.

`first()` and `last()` can be nested.

```{r}
read.table(header=TRUE, text ="
           Temp.Max Temp.Mean Temp.Min
2016-07-01       74        69       60
2016-07-02       78        66       56
2016-07-03       79        68       59
2016-07-04       80        76       69
2016-07-05       90        79       68
2016-07-06       89        79       70
2016-07-07       87        78       72
2016-07-08       89        80       72
2016-07-09       81        73       67
2016-07-10       83        72       64
2016-07-11       93        81       69
2016-07-12       89        82       77
2016-07-13       86        78       68
2016-07-14       89        80       68
2016-07-15       75        72       60
2016-07-16       79        69       60") -> t

temps <- as.xts(t)

# Create lastweek using the last 1 week of temps
lastweek <- last(temps, "1 week")

# Print the last 2 observations in lastweek
last(lastweek, n=2)

# Extract all but the first two days of lastweek
first(lastweek, "-2 days")
```

### Compound `first()` & `last()`

```{r}
# Extract the first three days of the second week of temps
first(last(first(temps, "2 weeks"), "1 week"), "3 days")
```

## Math operations

### Key faatures

* xts is naturally a matrix
  * Most operations are standard matrix
* **Math operations are on the _intersection_ of times**
  * 2 objects will first be merged on the time index then the operation applied
* Sometimes it's necessary to drop the xts class
  * argument drop = TRUE, coredata(), or as.numeric()
* Special handling required for _union_ of dates

### Operations on unions
* May be necessary to use all observations

### Example

```{r}
a <- as.xts(read.table(header=TRUE, text="
           a
2015-01-24 1
2015-01-25 1
2015-01-26 1"))

b <- as.xts(read.table(header=TRUE, text="
           b
2015-01-24 2"))

print(a + b)

a + as.numeric(b)
```

### Merging example

Merging is covered in Chapter 3 but here's an example of what's possible.

```{r}
# Add a to b, and fill all missing rows of b with 0
a + merge(b, index(a), fill = 0)

# Add a to b and fill NAs with the last observation
a + merge(b, index(a), fill = na.locf)
```



