---
title: "Data Camp / xts - Chapter 1"
output:
  pdf_document: default
  html_notebook: default
---

Chapter 1 of Data Camp xts class excersizes

xts objects can be thought of as a matrix with a corresponding time component.

```{r}
library(xts)

# Create the object data using 5 random numbers
data = rnorm(5)

# Create dates as a Date class object starting from 2016-01-01
dates <- seq(as.Date("2016-01-01"), length = 5, by = "days")

# Use xts() to create smith
smith <- xts(x = data, order.by = dates)

# Create bday (1899-05-08) using a POSIXct date class object
bday <- as.POSIXct("1899-05-08")

# Create hayek and add a new attribute called born
hayek <- xts(x = data, order.by = dates, born = bday)

print(data)
print(dates)
print(smith)
print(bday)
print(hayek)
print(str(hayek))

```


## Getting at the data

### From zoo

`coredata()` is used to get the core matrix.  
`index()` is used to get the index.

```{r}
# Extract the core data of hayek
hayek_core = coredata(hayek)

# View the class of hayek_core
class(hayek_core)

# Extract the index of hayek
hayek_index = index(hayek)

# View the class of hayek_index
class(hayek_index)
```

xtc will convert any valid date format to an internal format that makes indexing the data easy.

The _row names_ **are** the time values.

```{r}
a <- xts(x = 1:2, as.Date("2012-01-01") + 0:1)
a[index(a)]
```

When subsetting, any valid date type can be used.  However, you must be careful to make sure compatable
date types are used.  For example, a POSIX can index an xtc with an index of Date type but Date types can
not be used directly to index an xtc with POSIX index, it has to be converted to POSIX first.

```{r}
# Create dates
dates <- as.Date("2016-01-01") + 0:4

# Create ts_a
ts_a <- xts(x = 1:5, order.by = dates)

# Create ts_b
ts_b <- xts(x = 1:5, order.by = as.POSIXct(dates))

# Extract the rows of ts_a using the index of ts_b
ts_a[index(ts_b)]

# Extract the rows of ts_b using the index of ts_a
ts_b[as.POSIXct(index(ts_a))]
```

## Reading / Writing Converting

### Converting

Like most r classes, xts supports the as.* approach.

```{r}
# Convert austres to an xts object called au
au <- as.xts(austres)

# Then convert your xts object (au) into a matrix am
am <- as.matrix(au)

# Inspect the head of am
head(am)

# Convert the original austres into a matrix am2
am2 <- as.matrix(austres)

# Inspect the head of am2
head(am2)
```

### Reading data

The standard read functions can be used to read in data as well as `read.zoo()`.

```{r}
# Create dat by reading tmp_file
dat <- read.csv('data/data_1.csv')

# Convert dat into xts
dat_xts_a <- xts(dat, order.by = as.Date(rownames(dat), "%m/%d/%Y"))

# Read tmp_file using read.zoo
dat_zoo <- read.zoo('data/data_1.csv', index.column = 0, sep = ",", format = "%m/%d/%Y")

# Convert dat_zoo to xts
dat_xts <- as.xts(dat_zoo)

print(dat_xts_a)
print(dat)
print(dat_zoo)
print(dat_xts)
```


### Exporting xts objects

Can use `write.zoo()` although it is faster to read `saveRDS()` / `writeRDS()`.

```{r}
# Convert sunspots to xts using as.xts().
sunspots_xts <- as.xts(sunspots)

# Get the temporary file name
tmp <- tempfile()

# Write the xts object using zoo to tmp 
write.zoo(sunspots_xts, sep = ",", file = tmp)

# Read the tmp file. FUN = as.yearmon converts strings such as Jan 1749 into a proper time class
sun <- read.zoo(tmp, sep = ",", FUN = as.yearmon)

# Convert sun into xts. Save this as sun_xts
sun_xts = as.xts(sun)

print(sunspots_xts)
print(sun)
print(sun_xts)
```

