---
title: "Data Camp / Forecasting Using R - Chapter 5"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Initialize and load the data

```{r}
library(ggplot2)
library(forecast)
library(fpp)
```

Chapter 5 of [Data Camp - Forecasting Using R](https://www.datacamp.com/courses/forecasting-using-r)

## Dynamic Regression

Including **external** variables.  It is similar to a **regression** but the _error term_ is an ARIMA process.

$y_t = \beta_0 + \beta_1 x_{1,t}+ ... + \beta_r x_{r,t} + e_t$

The **error term** is an ARIMA process, vs. normal Regression where it is _white noise_.

The **ARIMA** process is where the historical informatino about the time series is incorporated.

It is modeled with the `auto.arima()` function but with an additional argument, **xreg=** which takes a matrix of predictor variables to include in the model.

The forecast will require forecasted predictor values as well.  This is done wiht the **xreg=** argument which takes a matrix of the predictor values.

## Dyname harmonic regression

Periodic seasonality can be handled using pairs of Fourier terms:

$s_k(t) = sin(\frac{2\pi kt}{m})$  $c_k(t) = cos(\frac{2\pi kt}{m})$

$y_t = \beta_0 + \sum_{K=1}^K[\alpha_K s_k(t) + \gamma_K c_K(t)] + e_t$

* m = seasonal period
* Every periodc function can be approximated in terms of sin and cos terms for large enough $K$
* $\alpha_k$ and $\gamma_k$ are coefficients
* $e_t$ can be modeled as a non-seasonal ARIMA process

This method, using the Fourier terms, assumes the seasonal pattern does **NOT** change over time, whereas a seasonal ARIMA can handle models with changing seasonality.

Below will performs a log transformation, **labmda=0**, indicates to **NOT** do a seasonal ARIMA on the errors.

```{r}
fit = auto.arima(cafe, xreg=fourier(cafe, K=1), seasonal=FALSE, lambda=0)
fit %>% forecast(xreg=fourier(cafe, K=1, h=24)) %>%
  autoplot() + ylim(1.6, 5.1)
```


$y_t =\beta_0 + \beta_1 x_{t,1} + ... + \beta_{t,r} x_{t,r} + \sum_{K=1}^K[\alpha_K s_k(t) + \gamma_K c_K(t)] + e_t$

* Choose K to minimize AICc
* K can not be more than m/2

Using Fourier approach can handle very large sesonal periods.

## TBATS model

* Trigonometric terms for seasonality
* Box-Cox transformations for heterogeneity
* ARIMA errors for short-term dynamics
* Trend (possibly damped)
* Seasonal (including multiple and non-integer periods)

And best of all, they are automatgic!  **WARNING**: This can be dangerous as they do not always pick the best values.

TBATS is good for non-integer seasonality and multiple seasonal periods **BUT** does not always do a good job of predicting.