---
title: "Data Camp / Forecasting Using R - Chapter 3"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Initialize and load the data

```{r}
library(ggplot2)
library(forecast)
library(fpp)
```

Chapter 3 of [Data Camp - Forecasting Using R](https://www.datacamp.com/courses/forecasting-using-r)

## Exponential Forecast

### Simple Exponential smoothing

In R uses the `ses()` function.  The parameters (noted below) are estimated using least squares estimation.

**Forecasting Notation:**

$\hat{y}_{t+h|t}=$ point forecast of $y_{t+h}$ given data $y_1, ..., y_t$

**Forecast Equation:**

$\hat{y}_{t+h|t} = \alpha y_t + \alpha(1 - \alpha) y_{t-1} + \alpha(1 - \alpha)^2 y_{t-2} + ...$ Where $0 \leq \alpha \leq 1$

$\alpha$ determines how much weight is applied to more recent observations and how quickly they trail away.  Larger values means that more weight on recent observation(s) and decay more slowly.

This can be (more conveniently) represented as:

Forecast equation:   $\hat{y}_{t+h|t} = \ell_t$

Smoothing equation:  $\ell_t = \alpha y_t + (1-\alpha)\ell_{t-1}$

* $\ell_t$ is the **level** (smoothed value) of the series at time t.

The forecast is the most recent estimate of the $\ell$ **(level)**.

$\ell$ _evolves_ over time based on the most recent observation and the previous estimate of the level.

There are 2 components to estimate $\alpha$, initial value of $\ell_0$

This is all handled in R by the `ses()` function.

```{r}
data(oil)
oildata = window(oil, start=1996)
fc = ses(oildata, h=5)
summary(fc)
```


What does the above tell us?

alpha tells us that 79% of the forecasting is done on the first observation, then 

((1 - .79) * .79)% on the second and the remaining on the remaining points.  $\ell_0$ is 448.1323.

h was set to 5 to take us out the next 5 years.  This forecast is stored in the fc object.

```{r}
autoplot(fc) +
  ylab("Oil (millions of tonnes)") +
  xlab("Year")
```

## Holt's linear trend

You need to add to the simple exponential smoothing a component for trending.  This leaves us with the following equasions:

Forecast equation:   $\hat{y}_{t+h|t} = \ell_t + hb_t$

Level (Smoothing equation):  $\ell_t = \alpha y_t + (1-\alpha)(\ell_{t-1} + b_{t-1})$

Trend : $b_t = \beta^*(\ell_t - \ell_{t-1}) + (1 - \beta^*)b_{t-1}$

Because the slope can change this is often called a "Local Linear Trend"

Smaller $\beta^*$ means slower slope changes.  
Larger $\beta^*$ means faster slope changes.

There are now **4** parameters

* 2 for smoothing $\alpha$ and $\beta^* (0 \leq \alpha,\beta^* \leq 1)$
* 2 state parameters $\ell_0$ and $b_0$
* Choose $\alpha, \beta^*, \ell_0, b_0 to minimize SSE$

The **Holt** method will have a constant slope over time.

The `holt()` function handles the minimizing of parameters.

Here is an example of forecasting with Holt's method in R, forecasting for 5 years forward, using data from 1990 - 2004.



```{r}
ausair %>% 
  window(start=1990, end=2004) %>%
  holt(h=5) %>%
  autoplot
```

## Damped trend method

The Dampled trend method will allow the slope to dampen over time.  The equasions for this are:

$\hat{y}_{t+h|t} = \ell_t + (\phi + \phi^2 + ... + \phi^h)b_t$  
$\ell_t = \alpha y_t + (1 - \alpha)(\ell_{t-1} + \phi b_{t-1})$
$b_t = \beta^*(\ell_t - \ell_{t-1}) + (1-\beta^*)\phi b_{t-1}$

This is basically Holt's method with 1 extra parameter $\phi$ as a dampening parameter where $0 \leq \phi \leq 1$.  Where if $\phi = 1$ you have a linear trend identical to Holt's.

With this, the **short-run forecasts** are _trended_ but the **long-run forecasts** are _constant_.

**NOTE**: The **PI** parameter controls the upper/lower control bounds plotting. When FALSE, it's not displayed, when TRUE it is.  This is not documented in the help for `holt()`.

To add **dampening** add the **damped=TRUE** option to the `holt()` function call.

```{r}
window(ausair, start=1990, end=2004) -> airpassengers

holt(airpassengers, h=15, PI=FALSE) -> fc1
holt(airpassengers, h=15, PI=FALSE, damped=TRUE) -> fc2

autoplot(airpassengers) +
  xlab("Year") +
  ylab("millions") + 
  autolayer(fc1, series="Linear trend") +
  autolayer(fc2, series="Damped trend")
```

Below looks at visitors to austraila.

Using the Holt function to forecast 10 years out.

Note the checkrisudals has a high p-value so the residuals are not well correlated.  That's good!  
The ACF chart shows all values are insignficant, the residuals "look like" white noise and are relatively normally distributed so these all point to a "good" model.  Or, at least a model that isn't missing any information.

```{r}
holt(austa, h=10) -> fc
summary(fc)
autoplot(fc)
checkresiduals(fc)
```

## Exponential Smoothing methods with trenad and seasonality
### Holt - Winters

Holt - Winters works when there is both **trend** and **seasonality**
2 versions, additive and multiplicative.

The multaplicative method should be used if the seasonal variation increases with the level of the data.  This case it doesn't reall matter though.

```{r}
window(austourists, start=2005) -> aust

hw(aust, seasonal="additive") -> fc1
hw(aust, seasonal="multiplicative") -> fc2

autoplot(austourists) +
  autolayer(fc1, PI=FALSE, series="Additive") + 
  autolayer(fc2, PI=FALSE, series="Multiplicative")
```

## Taxonomy of exponential smoothing methods
<table border=1 black solid>
<tr><th></th><th colspan=3>Seasonal Component</th></tr>
<tr><th>Trend Component</th><th>N (None)</th><th>A (Additive)</th><th>M (Multiplicative)</th><tr>
<tr><th>N (None)</th><td>(N, N)</td><td>(N, A)</td><td>(N, M)</td></tr>
<tr><th>A (Additive)</td><td>(A, N)</td><td>(A, A)</td><td>(A, M)</td></tr>
<tr><th>Ad (Additive Damped)</td><td>(Ad, N)</td><td>(Ad, N)</td><td>(Ad, N)</td></tr>
</table>
<table border=1 black solid>
<tr><td>(N, N)</td><td>Simple exponential smoothing</td><td> ses()</td></tr>
<tr><td>(A, N)</td><td> Holt's linear method</td><td> holt()</td></tr>
<tr><td>(Ad, N) Additive damped trend method</td><td> hw()</td></tr>
<tr><td>(A, A)</td><td> Additive Holt-Winter's method</td><td> hw()</td></tr>
<tr><td>(A, M)</td><td> Multiplicative Holt-Winter's method</td><td> hw()</td></tr>
<tr><td>(Ad, M)</td><td> Damped multiplicative Holt-Winter's method</td><td> hw()</td></tr>
</table>


```{r}

# Plot the data
autoplot(a10)

# Produce 3 year forecasts
fc1 <- hw(a10, h=36, seasonal='multiplicative')
fc2 <- hw(a10, h=36, seasonal='additive')

# Plot forecasts
autoplot(a10) +
  autolayer(fc1, series='Multiplicative') +
  autolayer(fc2, series='Additive')
```

## Innovations state space models

* Each exponential smoothing method can be written as an **"innovations state space model"**
* Trend = ${N, A, A_d}$
* Seasonal  = ${N, A, M}$
* ERROR = ${A, M}$

_3 x 3 = 9 possible exponential smoothing methods_

## ETS models

**E**rror, **T**rend, **S**easonality models

* Parameters: estimated using the "likelihoo", the probability of the data arising from the specified model
* For models with additive errors, this is **equivolent to minimizing SSE**
* Choose the best model by minimizing a corrected version of Akaike's Information Criterion (AIC).  Similar to cross validation but much faster.

The `ets()` function does this for you!

for ausair the best model is **ETS(M, A, N)**.  This is a **M**ultiplicative errors, **A**dditive trend and **N**o seasonality.

A difference between `ets()` and `hw()` is that it doesn't create the forecast, you need to take the output and put it into the `forecast()` function and plot that.

We see in the example below the linear prediction.  And how the error grows over time from the multiplicative errors component.

```{r}
ets(ausair) -> fc
print(fc)
autoplot(fc)
fc %>% forecast() %>% autoplot()
```

A seasonal example

The model **ETS(M,Ad,M)** indicates **M**ultiplicative error, **A**dditive **d**ampening trend, **M**ultiplicative seasonality.

**NOTE**: ETS(<error>, <trend>, <seasonality>)

Note that because of the type of model there are many more parameters to estimate vs. the previous.

```{r}
ets(h02)
ets(h02) %>% forecast() %>% autoplot()
```

## When ETS goes bad (Fails)

```{r}
autoplot(lynx)

ets(lynx) -> fit

summary(fit)

fit %>% forecast(h=20) %>% autoplot()
```

