---
title: "Data Camp / Intro to Time Series Analysis - Chapter 5"
output:
  pdf_document: default
  html_notebook: default
---

Initialize and load the data

```{r}

```

Chapter 5 of [Data Camp - Intro to Time Series Analsys](https://www.datacamp.com/courses/introduction-to-time-series-analysis)

## Simple Moving Average Model

_Today = Mean + noise + Slope * (Yesterday's Noise)_

$Y_t = \mu + \epsilon_t + \Theta\epsilon_{t-1}$

Where $\epsilon_t$ is mean 0 WN.

3 parms:

* $\mu$
* Slope $\Theta$
* WN Variance 

If slope $\Theta$ is 0 then it's simply a WN model.  Otherwise $Y_t$ depends on both $\epsilon_t$ and $\epsilon_{t-1}$ and the process ${Y_t}$ is autocorrelated.

Large values of $\Theta$ lead to greater autocorrelation.

Netative values result in oscillatory ts.

```{r}
x <- arima.sim(model = list(ma=0.5), n = 100)

y <- arima.sim(model=list(ma=0.9), n=100)

z <- arima.sim(model=list(ma=(-0.5)), n=100)

plot.ts(cbind(x, y, z))

par(mfrow=c(3,1))
acf(x)
acf(y)
acf(z)
```

### MA Model Esitmation and Forecasting

Apply process to inflation from Ecdat package

```{r}
data(Mishkin, package='Ecdat')

inflation <- as.ts(Mishkin[,1])
inflation_changes <- diff(inflation)
par(mfrow=c(2,1))
ts.plot(inflation)
ts.plot(inflation_changes)

```

```{r}
par(mfrow=c(1,2))
ts.plot(inflation_changes)
acf(inflation_changes,lag.max=24)
```

The ACF shows a strong negative correlation at lag=1 and the remaining fluctuation near 0.

**ARIMA(0, 0, 0)** is a "First order" MA model.

```{r}
MA_inflation_changes = arima(inflation_changes, order=c(0, 0, 1))
print(MA_inflation_changes)
```

### Plot the fitted values

```{r}
ts.plot(inflation_changes)
MA_inflation_changes_fitted = inflation_changes - residuals(MA_inflation_changes)
points(MA_inflation_changes_fitted, type='l', col='red', lty=2)
```

### Use `predict()` to see the future!

```{r}
predict(MA_inflation_changes)

predict(MA_inflation_changes, n.ahead = 6)
```

# fit to Nile data

```{r}
MA <- arima(Nile, order = c(0, 0, 1))
print(MA)

ts.plot(Nile)
MA_fit <- Nile - resid(MA)
points(MA_fit, type = "l", col = 2, lty = 2)
```

### MA and AR process

MA and AR models are similar except that an MA model has autocorrelation at 1 lag and an AR model has autocorrelation at many lags.

**I**nformation **C**riteria are a way to check the goodness of fit.  Generally, lower values are better.  There are 2 available `AIC()`  **Akaike Information Criterion** and `BIC()` **Bayesian Information Criterion**.

