---
title: "Data Camp / Intro to Time Series Analysis - Chapter 3"
output:
  pdf_document: default
  html_notebook: default
---

Initialize and load the data

```{r}

```

Chapter 3 of [Data Camp - Intro to Time Series Analsys](https://www.datacamp.com/courses/introduction-to-time-series-analysis)

## Scatterplots

### Log returns

apply `diff(log(stock_price))` to each stock, then bind them with `cbind()` and plot.

_Log Returns_ or _continuously compounded returns_ are a commonly used value in financial time series analysis.  They are to log of gross returns.

```{r}
data("EuStockMarkets")

plot(EuStockMarkets)

# Use this code to convert prices to returns
returns <- EuStockMarkets[-1,] / EuStockMarkets[-1860,] - 1

# Convert returns to ts
returns <- ts(returns, start = c(1991, 130), frequency = 260)

# Plot returns
plot(returns)

# Use this code to convert prices to log returns
logreturns <- diff(log(EuStockMarkets))

# Plot logreturns
plot(logreturns)
```


### Calculating column stats

See the `colMeans()`, `colSum()`, etc or the dot versions, `.colMeans()` for calculating column or row statistics on a dataset.

### Covariance

`cov()` is used to calculate the _covariance_ between to datasets which are related over time.  Positive covariance implies a positive relationship.  _Covariance_ measures teh strength of the linear relationship between matched pairs of variables.

_covariance_ are dependant on the size of the values. So it's hard to interpret the value.

### Correlations

* Standardized version of _covariance_ between -1 and 1.
* +1 is perfect positive linear relationship
* -1 is perfect negative linear relationship
* - is no linear relationship

`cor()` is used to calculate the _correlation_.  This is the _covariance_ divided by the product of the _sd_ of the 2 inputs.  e.g.  `cov(stock_a, stock_b) / (sd(stock_a) * sd(stock_b))`.

`cor()` and `cov()` can be used on 2 time series or on a matrix with multiple colums to compare.

```{r}
DAX_logreturns <- ts(diff(log(EuStockMarkets[,'DAX'])), start=c(1991, 131), frequency=260)
FTSE_logreturns <- ts(diff(log(EuStockMarkets[,'FTSE'])), start=c(1991, 131), frequency=260)

cov(DAX_logreturns, FTSE_logreturns)

# Use cov() with logreturns
cov(logreturns)

# Use cor() with DAX_logreturns and FTSE_logreturns
cor(DAX_logreturns, FTSE_logreturns)

# Use cor() with logreturns
cor(logreturns)
```

### Autocorrelation

A means to determine how a value relates to a previous value.

Autocorelation at lag 1 is comparing current value to the previous, lag=2 todays prior.

### The AutoCorrelation Function (ACF)

The `acf()` function calculates the Autocorrelation for a series based on the indicated lag.  It will also plot the autocorrelation(s) if _plot=TRUE_ is set.  If no lag is specified, it will calculate 0 - 10.

**NOTE:** acf calculates the correlation using 1/n, which is preferable for time series, wherase the `cor()` function uses 1/n-1.  This means that the `acf()` value is close but off by a factor of n-1/n.

```{r}
x = ts(arima.sim(model=list(order=c(0,0,0)), n=150),
        start=1, frequency=1)

x0 = x[-1]
x1 = x[-150]

head(cbind(x0, x1))

plot(x0, x1)

cor(x0, x1)

acf(x, lag.max = 1, plot=FALSE)

cor(x0, x1) * (149 / 150)
```


