---
title: "Data Camp / Intro to Time Series Analysis - Chapter 1"
output:
  pdf_document: default
  html_notebook: default
---

Initialize and load the data

```{r}
library(datasets)
data(AirPassangers)
```

Chapter 1 of [Data Camp - Intro to Time Series Analsys](https://www.datacamp.com/courses/introduction-to-time-series-analysis)

**Def**: A time series is a sequence of data in chronological order

`print()` prints the data with time stamps and includes the start, end and frequency informatino.

`plot()` plots a time series with time on the horizontal axis.

The course will cover basic time series models

* White Noise (WN)
* Random Walk (RW)
* Autoregression (AR)
* Simple Moving Average (MA)

### Examples

```{r}
data(Nile)

class(Nile)

str(Nile)

print(Nile)

length(Nile)

head(Nile, n=5)
tail(Nile, n=5)
```

The `plot()` function "knows" about time series and will plot them appropriately. Here are some examples

```{r fig.height=10}
par(mfrow=c(3, 1))

plot(Nile)
plot(Nile, 
     xlab = "Year", 
     ylab = "River Volume (1e9 m^{3})")

plot(Nile, 
     xlab="Year", 
     ylab="River Volumne (1e9 m^{3})", 
     main="Annual River Nile Volume at Aswan, 1871-1970",
     type='b')
```

The **ts** version of `plot()` provides for a _y=_ value which will be used as the "y" axis values.

When plotting a _Continuous time index_ the points will not be evenly spaced, they will be spaced based on the
_continuous_ value.  To addresss this, you can create a _discrete_ index.

```{r}
cti = 1:20 + runif(20)
cts = c(rnorm(10), rnorm(10))
dti = 1:20

par(mfrow=c(2, 1))

plot(cti, cts, type='b')
plot(dti, cts, type='b')
```

### Sampling Frequency

Time series data can be equally spaced, such as hourly temperature measurements, or unevently spaced such as the temperature every time you check e-mail.  Time series can also have _missing_ values.  For example, stocks are not traded on weekends, so there will be missing daily data for Saturdays and Sundays as well as possible holidays.

There are some _assumptions_ made about a time series.

* Consecutive obs are equally spaced
* The have a discrete-time observation index
* This may hold _approximately_ as in the case of stock trading with weekends and holidays missing.

4 Functions can be used to validate a time series.

* `start()`, which will return the start values, i.e. 1 1 may indicate it starts on hour 1 of day 1
* `end()`, returns the end values.  i.e. 1 24 may indicate it ends on hour 24 of day 1
* `frequency()` returns the frequency, how many observations.  i.e. 24
* `deltat()` returns the time interval duration.  i.e. 0.0417 is 1/24 or hourly.
* `cycle()` returns the position of an obs in the cycle of each obs.

```{r}
plot(AirPassengers)
print("Start")
start(AirPassengers)
print("end")
end(AirPassengers)

print("time")
time(AirPassengers)
print("deltat")
deltat(AirPassengers)
print("frequency")
frequency(AirPassengers)
print("cycle")
cycle(AirPassengers)
```

### Missing values.

Missingness is represented by NA in R.  There are many way's to handle it and differnet functions in R handle it differently.  Missing data can be removed or imputed (calculated) from the remaining data.

```{r}
data("AirPassengers")
AirPassengers[seq((7*12)+1, length.out=12)] <- NA  # NA out 19576

par(mfrow=c(2, 1))
plot(AirPassengers)

AirPassengers[85:96] <- mean(AirPassengers, na.rm = TRUE)

plot(AirPassengers)

data("AirPassengers")
points(AirPassengers, type = "l", col = 2, lty = 3)
```

### The **ts** class

To turn something into a time series use the `ts()` function.

```{r}
data_vector = rnorm(50)

head(data_vector)
par(mfrow=c(2, 1))

plot(data_vector)

time_series = ts(data_vector, start=2004, frequency=4)   # start in 2004 with quarterly frequency

print(time_series)
plot(time_series)
```

### Check if it's a _ts_

`is.ts()` returns TRUE if an object is a _ts_ and FALSE if not.

```{r}
data("EuStockMarkets")

is.ts(EuStockMarkets)

start(EuStockMarkets)
end(EuStockMarkets)
frequency(EuStockMarkets)

plot(EuStockMarkets)

ts.plot(EuStockMarkets, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")
legend("topleft", colnames(EuStockMarkets), lty = 1, col = 1:4, bty = "n")
```






