---
title: "Data Camp / Intro to Time Series Analysis - Chapter 2"
output:
  pdf_document: default
  html_notebook: default
---

Initialize and load the data

```{r}

```

Chapter 2 of [Data Camp - Intro to Time Series Analsys](https://www.datacamp.com/courses/introduction-to-time-series-analysis)

### Trends: Variance

#### log()

* Can _linearize_ rapid growth
* Can stabalize increasing variance

**Only defined for positive values**

#### diff()

Can remove the linear trend

With the _seasonal difference option_ can remove the seasonal component `diff(..., s)`

```{r}
rapid_growth = 500 + (seq(1, 40, length.out = 200) ^ 2) + (runif(200) * seq(1, 400, length.out=200))

par(mfrow=c(2,1))

ts.plot(rapid_growth)

linear_growth = log(rapid_growth)

ts.plot(linear_growth)
```

`diff()` takes the difference between each adjacent value.  The result is always _n -1_ records.  `diff()` can remove a time trend from a series. 

Let's see what happens when we apply to the linear_growth data.

```{r}

diff_linear_growth = diff(linear_growth)
par(mfrow=c(2,1))
ts.plot(linear_growth)
ts.plot(diff_linear_growth)
```

For data with seasonality the _lag=_ option can be applied to remove the seasonality.  Let's see what this can do.

```{r}
rapid_growth_with_seasonality = rapid_growth + sin(1:200) * 100

par(mfrow=c(2, 1))
# smooth it
linear_growth_with_seasonality = log(rapid_growth_with_seasonality)
linear_growth_with_seasonality_no_trend = diff(linear_growth_with_seasonality)
ts.plot(rapid_growth_with_seasonality)
ts.plot(linear_growth_with_seasonality)
ts.plot(linear_growth_with_seasonality_no_trend)
ts.plot(diff(linear_growth_with_seasonality_no_trend, lag=6))
```

### White Noise model

Simple examle of a _stationary process_.

The _white noise process_ has:

* A fixed, constant mean
* A fixed, constant variance
* No correlation over time

If any of these and possibly others, are true it is **NOT** _WN_

* There is a trend in the data
* There is periodocity in the data.
* The variance is changing over time

The `arima.sim()` function can simulate a _WN_ dataset.  `arima.sim(model=list(order=c(0,0,0)), n=50)`.
#### ARIMA

**A**uto**R**egressive **I**ntegrated **M**oving **A**verage

The `arima()` function will display information about an _ARIMA_ dataset including the _mean_ (_intercept_), _variance_, _standard error_, etc.

These values can also be obtained directly with the `mean()` and `var()` functions.

```{r}
wn1 = arima.sim(model = list(order=c(0,0,0)), n=50)
wn2 = arima.sim(model = list(order=c(0,0,0)), n=50, mean=100, sd=10)

par(mfrow=c(2,1))
ts.plot(wn1)
ts.plot(wn2)

arima(wn1, order=c(0,0,0))
mean(wn1)
var(wn1)
sqrt(var(wn1))
arima(wn2, order=c(0,0,0))
```

### Random Walk Time Series

* No specified mean or variance
* Strong dependance over time
* Change or ingrements are WN

The _Random Walk Model_ is recursively defined as:

_Today = Yesteray + Noise_

$Y_t = Y_{t-1} + \epsilon_t$

where $\epsilon_t$ is mean zero white noise (WN) and $\epsilon_t \longrightarrow \text{diff(Y)}$ is WN

* Simulation requires and initial point $Y_0$
* It has only 1 parm, the WN variance $\sigma_{\epsilon}^2$

The `diff()` of a _Random Walk_ series is a _White Noise_ series.

It can be defined to include a _drift_ variable represented by $c$

$Y_t = c + Y_{t-1} + \epsilon_t$

A _drift_ model has 2 parms, the WN variance $\sigma_{\epsilon}^2$ and the constant $c$.  Which is just a _Random Walk_ series with WN with mean $c$.

The _Random Walk_ series is teh cumulative sum (or integration) of a mean zero WN series.  This is an **ARIMA(0, 1, 0)** model, where the middle 1 indicates the model's order of integration is 1.

`arima.sim()` is used to create a _Random Walk_ series with _order=c(0,1,0)_.

```{r}
random_walk = arima.sim(model=list(order=c(0,1,0)), n=100)

par(mfrow=c(1,2))

ts.plot(random_walk)

random_walk_diff = diff(random_walk)

ts.plot(random_walk_diff)
```

#### Now with _drift_

```{r}
rwd = arima.sim(model=list(order=c(0,1,0)), n=100, mean=1)
par(mfrow=c(1,2))

ts.plot(rwd)

rwdd = diff(rwd)
ts.plot(rwdd)
```


```{r}
model_wn <- arima(rwdd, order=c(0,0,0))
int_wn = coef(model_wn)

ts.plot(rwd)
abline(0, int_wn)
```

### Stationarity

* Stationary models are parsimonious
* Have distributional stability over time


Observed time series:

* Fluctuate randomly
* But behave similarly from one time period to the next

**_See additional slides for more details_**

```{r}
# Use arima.sim() to generate WN data
white_noise <- arima.sim(model=list(order=c(0,0,0)), n=100)

# Use cumsum() to convert your WN data to RW
random_walk <- cumsum(white_noise)
  
# Use arima.sim() to generate WN drift data
wn_drift <- arima.sim(model=list(order=c(0,0,0)), n=100, mean=0.4)
  
# Use cumsum() to convert your WN drift data to RW
rw_drift <- cumsum(wn_drift)

# Plot all four data objects
plot.ts(cbind(white_noise, random_walk, wn_drift, rw_drift))
```




